---
title: "Lab2"
author: "Jason Chen"
date: "2024-08-29"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1
```{r}
  getwd()
```
# Task 2
```{r}
  epagas <- read.csv("EPAGAS.csv")
  head(epagas)
```
# Task 3
  Make the object mpg, the number of miles per gallon vector
```{r}
  mpg <-epagas$MPG
  head(mpg)
```
  Transform mpg variable to z and verify these results
```{r}
  z <- (mpg - mean(mpg)) / sd(mpg)
  head(z)
  print(paste0("mean = ", round(mean(z), digits = 4)), quote =               FALSE)
  print(paste0("variance = ", var(z)), quote = FALSE)
  print(paste0("standard deviation = ", sd(z)), quote = FALSE)
```
  Using z, find values of mpg that are possible outliers
```{r}
  mpg[abs(z)>=2 & abs(z)<=3]
```
  Using z, find the values of mpg that are defined as outliers
```{r}
  mpg[abs(z)>3]
```
  Using the lattice package, construct a dotplot with colors
```{r}
  library(lattice)
  mycol = ifelse(abs(z)>3, "Red",
        ifelse(abs(z)>=2 &abs(z)<=3,"Blue", "Black"))
  dotplot(z, col=mycol)
```
# Task 4
  Make a boxplot of the mpg variable
    - Make the box black
    - Put a notch where the median goes
    - Put a title on the graph
    - Make the plot horizontal
```{r}
  boxplot(mpg, main="MPG Values", xlab = "MPG", col = "BLACK", horizontal = TRUE, notch = TRUE)
```
  Chebyshev's Rule says that at least (1 - 1/k^2) of the              measurements will fall within k standard deviations of the mean.
```{r}
  k = 2
  cheb <- 1 - 1 / (k ^ 2)
  cheb
```
  Exact proportion within 2 standard deviations of the mean of mpg
```{r}
  exact <- length(mpg[abs(z < 2)]) / length(mpg)
  exact
```
  The Empirical Rule says that approximately 95% of the measurements   will lie within 2 standard deviations of the their mean given the dataset has an approximately mound-shaped, symmetric distribution.
  
  The Empirical Rule is very close because it states 95% will be      within 2 standard deviations while the exact proportion is 98%.
  
  The Empirical Rule is valid in this case because the dataset       looks symmetrical and mound-shaped.
```{r}
  densityplot(mpg)
```

